Enzo+Moray: Radiation Hydrodynamics Adaptive Mesh Refinement Simulations with Adaptive Ray Tracing

### 2010-09-08 ms.tex
Finished strong scaling section and added figure.  Need to add results without ray merging.

\subsection{Parallelization Strategy}

@@

Strong scaling tests shows how the problem scales with the number of
processors for the same problem.  The overhead associated with the
structured AMR framework in \enzo~can limit the strong scalability.
One key property of strong scaling is that each processor must have
sufficient work to compute, compared to the communication involved.
Here we use a small-box reionization calculation with $L_{\rm box} =
3$ Mpc/h, a resolution of $256^3$, and 6 levels of refinement.  We
measure the time spent on the hydrodynamics, non-equilibrium
chemistry, the rebuilding of the AMR hierarchy, and ray tracing in a
single level-0 timestep, lasting 5 Myr, at $z=9.5$.  There are ??? point
sources in this calculation at this redshift.  The ionized volume
fraction is ???.

>>>>

\subsection{Parallelization Strategy}
\label{sec:parallel}

@@

\begin{figure}[t]
  \epsscale{0.8}
  \plotone{fig/strong_v1.eps}
  \caption{\label{fig:strong} Strong scaling test with a $256^3$ AMR
    calculation of cosmological reionization.  The error bars
    represent the minimum and maximum time spent on a single core and
    measures load balancing.  The hydrodynamics and non-equilibrium
    chemistry solvers scale well.  Unfortunately the communication of
    rays does not scale in this problem, where as the ray tracing
    scales relatively well.}
  \epsscale{1}
\end{figure}

Strong scaling tests shows how the problem scales with the number of
processors for the same problem.  The overhead associated with the
structured AMR framework in \enzo~can limit the strong scalability.
One key property of strong scaling is that each processor must have
sufficient work to compute, compared to the communication involved.
Here we use a small-box reionization calculation with $L_{\rm box} =
3$ Mpc/h, a resolution of $256^3$, and six levels of refinement.  We
measure the time spent on the hydrodynamics, non-equilibrium
chemistry, ray tracing, and radiation transport communication in a
timestep lasting 1 Myr, at $z=10$.  There are nine radiative transfer
timesteps in this period.  The box has 675 point sources, ??? AMR
grids, and ??? computational cells in this calculation at this
redshift.  In the first radiative transfer timestep, ??? and ??? ray
segments are traced with and without ray merging.  The ionized volume
fraction is ???.  The calculations are performed on the Nehalam nodes
on Pleiades on 2$^n$ cores, where $n = [2,...,8]$.

Figure \ref{fig:strong} shows the strong scaling results of this
calculation.  The hydrodynamics and non-equilbrium chemistry routines
scale very well in this range because they depend on local phenomena.
Note that the dominant process is the radiation transport instead of
the hydrodynamics when compared to the weak scaling tests.  The error
bars in the figure represent the deviations across all cores, and it
shows that load balancing becomes an issue at 128 and 256 cores.  This
is an even larger problem for the ray tracing because the grids that
host multiple radiation sources will have significantly more work than
a distant grid.  For example if a few central grids host several point
sources, the exterior grids must wait until those grids are finished
ray tracing in order to receive the rays.  This idle time constitutes
the majority of the time spent in the ray communication routines
despite our efforts (see \S\ref{sec:parallel}) to minimize idle time.
For this reason, the communication of rays does not scale in this
problem.  One solution is to split the AMR grids into smaller blocks
based on the ray tracing work.  This could impact the performance of
the rest of Enzo by increasing the number of boundaries and thus
communication.  Nevertheless in simulations that are dominated by
radiation transport, it will be advantageous to construct such a
scheme to increase the feasbility of running larger problems with the
continually increasing computational resources.
